FMA: A Dataset For Music Analysis

### 2017-04-21 fma.tex
anonymous initial submission to ISMIR

\multauthor
{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}
{
	$^*$LTS2, EPFL, Lausanne, Switzerland \hspace{0.4cm}
	$^\dagger$SCSE, NTU, Singapore \hspace{0.4cm}
	{\small Work conducted when XB was at EPFL.}\\
	{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
  }

@@

%% To make customize author list in Creative Common license, uncomment and customize the next line
\def\authorname{Kirell Benzi, Michaël Defferrard, Pierre Vandergheynst, Xavier Bresson}

@@

% about 150-200 words.
\begin{abstract}
We introduce the Free Music Archive (FMA), an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography, tags, and comments; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 songs from 8 major genres, a medium genre-unbalanced set of 25,000 songs from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.

@@

\begin{table}[t]
	\centering
	\begin{threeparttable}
	\begin{tabular}{l@{ }rrcc}
		\toprule
		dataset\tnote{1} & \#clips & \#artists & year & audio \\
		\midrule
		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 729 & - & 2004 & yes \\
		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		% 103-Artists is also known as DB-L, Extended Ballroom
		\item[1] Names are clickable links to datasets' homepage.
		\item[2] Audio not directly available, can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
		\item[3] The 25,863 clips are cut from 5,405 songs.
		\item[4] Low quality 16 kHz, 32kbps, mono mp3.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison with some other datasets.} % Genre classification datasets.
	\label{tab:datasets}
\end{table}
% Lists of datasets:
% * http://www.audiocontentanalysis.org/data-sets/
% * A Survey of Evaluation in Music Genre Recognition

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & [days] \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Scale is the number of samples times the dimensionality. Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776

@@

We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
We are grateful to SWITCH and EPFL for hosting the dataset within the context of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC P-2 program}.
% We are grateful to \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
Xavier Bresson is supported by NRF Fellowship NRFF2017-10.

>>>>

\multauthor
{\phantom{Michaël Defferrard$^*$ \hspace{.6cm} Kirell Benzi$^*$ \hspace{.6cm} Pierre Vandergheynst$^*$ \hspace{.6cm} Xavier Bresson$^\dagger$}}
{\phantom{
	$^*$LTS2, EPFL, Lausanne, Switzerland \hspace{0.4cm}
	$^\dagger$SCSE, NTU, Singapore \hspace{0.4cm}
	{\small Work conducted when XB was at EPFL.}}\\
	\phantom{\tt\small \{michael.defferrard,kirell.benzi,pierre.vandergheynst\}@epfl.ch, xbresson@ntu.edu.sg}
  }

@@

%% To make customize author list in Creative Common license, uncomment and customize the next line
\def\authorname{\phantom{Kirell Benzi, Michaël Defferrard, Pierre Vandergheynst, Xavier Bresson}}

@@

% about 150-200 words.
\begin{abstract}
We introduce the Free Music Archive (FMA), an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography, tags, and comments; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 songs from 8 major genres, a medium genre-unbalanced set of 25,000 songs from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
Code, data, and usage examples are available at \url{https://github.com/anonymous}.

@@

\begin{table}[t]
	\centering
	\begin{threeparttable}
	\begin{tabular}{l@{ }rrcc}
		\toprule
		dataset\tnote{1} & \#clips & \#artists & year & audio \\
		\midrule
		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 729 & - & 2004 & yes \\
		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
		\bf \href{https://github.com/anonymous/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		% 103-Artists is also known as DB-L, Extended Ballroom
		\item[1] Names are clickable links to datasets' homepage.
		\item[2] Audio not directly available, can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
		\item[3] The 25,863 clips are cut from 5,405 songs.
		\item[4] Low quality 16 kHz, 32kbps, mono mp3.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison with some other datasets.} % Genre classification datasets.
	\label{tab:datasets}
\end{table}
% Lists of datasets:
% * http://www.audiocontentanalysis.org/data-sets/
% * A Survey of Evaluation in Music Genre Recognition

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & [days] \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
		\href{https://github.com/anonymous/}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Scale is the number of samples times the dimensionality. Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776

@@

\phantom{
We want to thank the team supporting the \href{https://freemusicarchive.org}{Free Music Archive} as well as all the contributing artists and curators for the fantastic content they made available.
We are grateful to SWITCH and EPFL for hosting the dataset within the context of the \href{https://projects.switch.ch/scale-up}{SCALE-UP} project, funded in part by the swissuniversities \href{http://www.swissuniversities.ch/isci}{SUC P-2 program}.
% We are grateful to \href{https://zenodo.org}{zenodo.org}, whose free hosting and archiving of digital research artifacts is promoting a more open scientific process.
Xavier Bresson is supported by NRF Fellowship NRFF2017-10.
}
